---
title: "data 605 wk2 assignment"
author: "Jun Pan"
date: "February 9, 2019"
output: html_document
---

1. Problem set 1
(1) Show that AT A 6= AAT
in general. (Proof and demonstration.)
#demonstration
```{r}
A <- matrix(c(1,2,1,4,0,-1,3,1,-2), nrow=3, byrow=TRUE)
At <- t(A)
```
# Consider matrix A
```{r}
A
```

# Its transpose At
```{r}
At
```
# Multiplying A by its transpose At
```{r}
A %*% At
```
# Multiplying transposed A by A 
```{r}
At %*% A
```
# Check using logical comparison
```{r}
A %*% At == At %*% A
```

(2) For a special type of square matrix A, we get AT A = AAT
. Under what conditions
could this be true? (Hint: The Identity matrix I is an example of such a matrix).
Please typeset your response using LaTeX mode in RStudio. If you do it in paper, please
either scan or take a picture of the work and submit it. Please ensure that your image is
legible and that your submissions are named using your first initial, last name, assignment
and problem set within the assignment. E.g. LFulton_Assignment2_PS1.png

#If a matrix is symmetrical along the diagonal. This is true because a symmetric matrix and its transpose are equal, meaning you are just multiplying a matrix by itself. If a matrix is symmetrical along the diagonal

#demonstration

```{r}
A <- matrix(c(1,3,0,3,1,0,0,0,1), nrow=3, byrow=TRUE)
```
# matrix A
```{r}
A
```
# its transpose is the same matrix
```{r}
t(A)
```

```{r}
A == t(A)
```

```{r}
(A %*% t(A)) == (t(A) %*% A)
```



2. Problem set 2
Matrix factorization is a very important problem. There are supercomputers built just
to do matrix factorizations. Every second you are on an airplane, matrices are being
factorized. Radars that track flights use a technique called Kalman filtering. At the heart
of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear
systems of equations when they track your flight using radars.
Write an R function to factorize a square matrix A into LU or LDU, whichever you
prefer. Please submit your response in an R Markdown document using our class naming
convention, E.g. LFulton_Assignment2_PS2.png
You don't have to worry about permuting rows of A and you can assume that A is less
than 5x5, if you need to hard-code any variables in your code. If you doing the entire
assignment in R, then please submit only one markdown document for both the problems.

```{r}
factorize_LU <- function(A) {
  # Check wheter matrix A is square
  if (dim(A)[1]!=dim(A)[2]) {
    return(NA)
  }
  
  U <- A
  n <- dim(A)[1]
  L <- diag(n)
  
  
  if (n==1) {
    return(list(L,U))
  }
  
  
  for(i in 2:n) {
    for(j in 1:(i-1)) {
      multiplier <- -U[i,j] / U[j,j]
      U[i, ] <- multiplier * U[j, ] + U[i, ]
      L[i,j] <- -multiplier
    }
  }
  return(list(L,U))
}
```

#Demonstration

```{r}
A <- matrix(c(2,1,1,1,1,3,7,3,4), nrow=3, byrow=TRUE)
LU <- factorize_LU(A)
L<-LU[[1]]  
U<-LU[[2]]

A
L
U
A == L %*% U
```

#### Example 2

```{r}
A <- matrix(c(1,3,4,7), nrow=2, byrow=TRUE)
LU <- factorize_LU(A)
L<-LU[[1]]  
U<-LU[[2]]

A
L
U
A == L %*% U
```
